Combinatorial interpretations of holonomic recurrences

### 2021-06-13 article/main.tex
Fix introduction

We analyze problem of finding combinatorial interpretations of holonomic recurrences and show some techniques that were used to achieve known results.

@@

In order to make reading this article easier, let's fix some notation and make some assumptions. Any deviation from this will be clearly stated.

\(\poly\) is set of all polynomials.

@@

Calligraphic \(\gf{X}\) is implicitly assumed to be generating function of the sequence \(X\) where \(X\) is any upper-case letter.

@@

\begin{definition}
    \label{def-holo-1}
    \cite{holotoolkit}
    An infinite sequence \(A = (a_i)_{i=0}^{\infty}\) is called holonomic (aka. P-finite, P-recursive or D-recursive) \(\textiff\) exists holonomic recurrence for it, ie.:

@@

\begin{example}
    Sequences \(a_n = \sqrt{n}\), \(b_n = n^n\) are \textbf{not} holonomic.
\end{example}

\begin{definition}
    \label{def-holo-2}
    \cite{holotoolkit, complexity}
    Equivalently to definition\ \ref{def-holo-1}, an analytic function \(y\) is called holonomic \(\textiff\) it satisfies differential equation in the following form:
    \[\sum_{i=0}^{r} p_i(z) y^{(i)}(z) = b(z)\]
    for some fixed \(r \in \N\); \(b, p_0, \ldots, p_r \in \poly\) not all identically zero.
\end{definition}

\begin{example}
    \(y(z) = \frac{5n - 3}{3n + 5}\) is holonomic, because:
    \[(5z-3)(3z+5)y'(z) - 34y(z) = 0\]
\end{example}

Equivalence of these two definitions assures us that if we find differential equation describing the desired objects satisfying definition\ \ref{def-holo-2}, we will be able to find holonomic recurrence mentioned in definition\ \ref{def-holo-1}.

@@

@@

>>>>

We analyze the problem of finding combinatorial interpretations of holonomic recurrences and show some techniques that were used to achieve known results.

@@

In order to make reading of this article easier, let's fix some notation and make some assumptions. Any deviation from this will be clearly stated.

\(\textiff\) means if and only if.

\(\poly\) is a set of all polynomials.

@@

Calligraphic \(\gf{X}\) is implicitly assumed to be a generating function of the sequence \(X\) where \(X\) is any upper-case letter.

@@

\begin{definition}
    \label{def-holo-1}
    \cite{holotoolkit}
    An infinite sequence \(A = (a_i)_{i=0}^{\infty}\) is holonomic (aka. P-finite, P-recursive or D-recursive) \(\textiff\) exists holonomic recurrence for it, ie.:

@@

\begin{example}
    Sequences \(a_n = \sqrt{n}\), \(b_n = n^n\) are \textbf{not} holonomic, ie. none of them satisifies a linear recurrence equation with polynomial coefficients\cite{nonholo}.
\end{example}

\begin{definition}
    \label{def-holo-2}
    \cite{complexity}
    Equivalently to definition\ \ref{def-holo-1}, \(A = (a_i)_{i=0}^{\infty}\) is holonomic \(\textiff\) exists generating function in the following form:
    \[\sum_{i=0}^{r} p_i(z) \gf{A}^{(i)}(z) = b(z)\]
    for some fixed \(r \in \N\); \(b, p_0, \ldots, p_r \in \poly\) not all identically zero.
\end{definition}

Equivalence of these two definitions assures us that if we find generating function describing the desired objects satisfying definition\ \ref{def-holo-2}, we will be able to find holonomic recurrence mentioned in definition\ \ref{def-holo-1}.\cite{complexity}

@@

    Our holonomic recurrence is then:
    \[\left\{\begin{array}{rcl}
                B(0) &=& 1\\
                (2 + 4 n) B(n) + (-2 - n) B(n + 1) &=& 0
    \end{array}\right.\]

    For our purposes we will often ignore the initial conditions, as they can be easily described by exact number of the objects. The most interesting part for us is the relation between the finite number consecutive elements.

@@

    \bibitem{nonholo}
    Stefan Gerhold,
    \textit{Combinatorial sequences: Non-holonomicity and inequalities. Ph.D. thesis},
    RISC-Linz, Johannes Kepler Universitat Linz,
    2005
